Neural networks for solving regression and classification problems

### 2022-11-15 doc/paper.tex
constand vs newton

In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function. IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON. Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT SCORE compared to the Newton method's HESSIAN SCORE.

@@

- GD vs SGD

Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2. Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds. Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.

An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization. As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time. With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization. For this specific plot, momentum does not seem to be an important parameter. Both the momentum and adagrad with momentum heatmaps are identical to their momentumless counterparts, constant and adagrad, implying that the best gridsearched momentum is 0. That momentum is irrelevant in this instance is not a generalizable result, as momentum has played a key role in previous plots and results.

Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same. As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adam, starts slowing down the rate at which the MSE is decreased. However, as can be seen in HEATMAP FIG, Adam offers a 11.11 \% \footnote{$\frac{constant MSE}{adam MSE} * 100 = \frac{0.018}{0.016} * 100 = 11.11\%$} decrease in MSE compared to the constant scheduler. It is clear from the results thus far that Adam is the best scheduler for approximating the Franke function when using linear regression with gradient descent.

Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE. Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE. Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model, in addition to other optimizations present in the scikit learn implementation. Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).  both falling short of the analytical MSE of 0,003. In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions. This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.

>>>>

In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features. We started off by comparing Newtons method for gradient descent to a constant learning rate using full batch gradient descent.

From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 within the first epoch. This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.  By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.21. Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script. Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}. This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex. However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations. For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly. Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.

@@

By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes. Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one. This means that our schedulers work the best for a fully stochastic gradient descent. The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of 
A full comparison of batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.

Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2. Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds. Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation. Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.

An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization. As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time. With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization. For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts. That being said, in FIG ALL it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.

Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same. As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased. However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler. It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent. We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs. The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d. We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.

To test this, we compared Adagrad with Adam over 5000 epochs. Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower. Scikit MSE 0.005.

Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE. Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE. We compared Adagrad to scikit learn's Adam MLPRegressor as this seemed to be the closest option for comparing our results. Even though we have cross validated both our own Adagrad implementation and Scikit's with 5 folds, these differences in results might be due to the difference in solver, a more favorable weight initiation in Scikits model or additional optimizations present in the scikit learn implementation. Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).  both falling short of the analytical MSE of 0,003. In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions. This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
